### 2025-07-31 Code/Paper_analysisandtables_R&R.Rmd
Clean out repository for manuscript publication getting rid of old versions of data and code

---
title: "Paper_analysisandtables"
author: "Kristin Dobbin"
date: "2024-09-18"
output: 
  html_document: default
  pdf_document: default
knit: (function(rmdfile, ...) { rmarkdown::render(rmdfile, output_dir=here::here("Docs"))})
---

```{r setup, include=FALSE, echo=FALSE}
#MARKDOWN SETTINGS
#knitr::opts_knit$set(root.dir = "/Users/KristinDobbin/Documents/Projects/R_Projects/Votingrights_systemperformance", warning=FALSE, message=FALSE, include=FALSE)

#LOAD LIBRARIES
library(tidyverse)
library(vtable)
library(sjPlot)
library(stargazer)

#LOAD AND READY DATA
Data <- read.csv(here::here("Data_processed/Systemperformancedata_RR.csv"))
Data <- Data[,-1]
Data$enfranchisement_final <- as.factor(Data$enfranchisement_final)
Data$Final_inst_update <- as.factor(Data$Final_inst_update)

#correct data types, levels and reference factors
Data[,c(1:11,16, 19:45, 47:61)] <- lapply(Data[,c(1:11,16, 19:45, 47:61)], factor)
levels(Data$WQ_combined_count) <- c(levels(Data$WQ_combined_count), "4")

Data$FAILING <- ifelse(Data$FINAL_SAFER_STATUS == "Failing", "Failing", "Not failing")
Data$FAILING <- as.factor(Data$FAILING)

Data$Source <- as.factor(Data$Source)
Data$public <- as.factor(Data$public)

#re-order factor levels for enfranchisement to be full, limited, none
Data$enfranchisement_final <- factor(Data$enfranchisement_final, levels = c("Full", "Limited", "None"))

Data$FAILING <- relevel(Data$FAILING, ref = "Not failing")
Data$enfranchisement_final <- relevel(Data$enfranchisement_final, ref = "Full")

#Make self produced water the reference (compared to purchased)
Data$Purchased <- relevel(Data$Purchased, ref = "Self-produced")

Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL <- relevel(Data$HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL, ref = "NONE")
Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL <- relevel(Data$TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL, ref = "NONE")
Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL <- relevel(Data$PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL, ref = "NONE")
Data$NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI <- relevel(Data$NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI, ref = "NONE")
Data$ABESENCE_OF_INTERTIES_RISK_LEVEL <- relevel(Data$ABESENCE_OF_INTERTIES_RISK_LEVEL, ref = "NONE")
Data$BOTTLED_WATER_OR_HAULED_WATER_RELIANCE_RISK_LEVEL <- relevel(Data$BOTTLED_WATER_OR_HAULED_WATER_RELIANCE_RISK_LEVEL, ref = "NONE")
Data$SOURCE_CAPACITY_VIOLATION_RISK_LEVEL <- relevel(Data$SOURCE_CAPACITY_VIOLATION_RISK_LEVEL, ref = "NONE")
Data$PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI <- relevel(Data$PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI, ref = "2")
Data$EXTREME_WATER_BILL_RISK_LEVEL_BI <- relevel(Data$EXTREME_WATER_BILL_RISK_LEVEL_BI, ref = "2")
Data$FUNDING_any_failingoratriskonly <- relevel(Data$FUNDING_any_failingoratriskonly, ref = "No")
Data$OPERATOR_CERTIFICATION_VIOLATIONS_RISK_LEVEL <- relevel(Data$OPERATOR_CERTIFICATION_VIOLATIONS_RISK_LEVEL, ref = "NONE")
Data$MONITORING_AND_REPORTING_VIOLATIONS_RISK_LEVEL <- relevel(Data$MONITORING_AND_REPORTING_VIOLATIONS_RISK_LEVEL, ref = "NONE")
Data$DAYS_CASH_ON_HAND_RISK_LEVEL <- relevel(Data$DAYS_CASH_ON_HAND_RISK_LEVEL, ref = "NONE")
Data$OPERATING_RATIO_RISK_LEVEL <- relevel(Data$OPERATING_RATIO_RISK_LEVEL, ref = "NONE")

#add in demographic data
Demographics <- read.csv(here::here("Data_processed/Systemdemographicdata.csv"))
Demographics <- Demographics[,-c(1,3)]
Demographics$PWSID <- as.factor(Demographics$PWSID)

Data <- left_join(Data, Demographics, by = "PWSID")

#Water quality
WQ1 <- glm(WQ_combined_any ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(WQ1)

##WQ sub components
ecoli <- glm(HISTORY_OF_E_COLI_PRESENCE_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(ecoli)

TT <- glm(TREATMENT_TECHNIUQE_VIOLATIONS_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(TT)

MCL <- glm(PERCENTAGE_OF_SOURCES_EXCEEDING_AN_MCL_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(MCL)

#Affordability 
Affordability1 <- glm(Affordability_combined_any ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(Affordability1)

## Affordability sub components
MHI <- glm(PERCENT_OF_MEDIAN_HOUSEHOLD_INCOME_MHI_RISK_LEVEL_BI ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(MHI)

extreme <- glm(EXTREME_WATER_BILL_RISK_LEVEL_BI ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(extreme)

covid <- glm(didnotapplycovid ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(covid)

# Accessibility
Accessibility1 <- glm(Accessibility_combined_any ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(Accessibility1)

## Accessibility sub components
Single_source <- glm(NUMBER_OF_WATER_SOURCES_RISK_LEVEL_BI ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(Single_source)

Sourcecapacity_violations <- glm(SOURCE_CAPACITY_VIOLATION_RISK_LEVEL ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(Sourcecapacity_violations)

Service_dis <- glm(Service_disruptions ~ enfranchisement_final + POPULATION + Source + Purchased + MHI + Percent.renter + Percent.white, data = Data, family= binomial)
summary(Service_dis)

```

This Rmd created html versions of the model results tables using stargazer. The models script must be run prior to using this Rmd

```{r table composites, results='asis', echo=FALSE}
#Composites
WQ1OR <- WQ1
WQ1OR$coefficients <- exp(WQ1OR$coefficients)
p.valuesWQ1 <- list(summary(WQ1)$coefficients[,4])

Affordability1OR <- Affordability1
Affordability1OR$coefficients <- exp(Affordability1OR$coefficients)
pvaluesAffordability1 <- list(summary(Affordability1)$coefficients[,4])

Accessibility1OR <- Accessibility1
Accessibility1OR$coefficients <- exp(Accessibility1OR$coefficients)
pvaluesAccessibility1 <- list(summary(Accessibility1)$coefficients[,4])

p.values2 <- list(c(1,1,1,1,1,1))

compositemodellist <- list(WQ1, WQ1OR, Affordability1, Affordability1OR,
                          Accessibility1, Accessibility1OR)

Compositestable <- capture.output(stargazer(compositemodellist,
                                 type = 'html', 
                                 column.labels = c("coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR"),
                                 model.numbers = FALSE, 
                                 dep.var.labels = c("Water quality", "Affordability", 
                                                    "Accessibility"), 
                                 ci=c(F,T, F, T, F, T), 
                                 star.cutoffs = c(0.05, 0.01, 0.001),
                                 notes="<span>&#42;&#42;&#42;</span>: p<0.001; <span>&#42;&#42;</span>: p<0.01;<span>&#42;</span>: p<0.05",
                                 notes.append = FALSE,
                                 p = c(p.valuesWQ1, p.values2, pvaluesAffordability1, p.values2,
                                       pvaluesAccessibility1, p.values2),
                                 omit.stat = c("ll"),
                                 covariate.labels = c("Limited enfran.", "No enfran.",
                                                      "Population", "Surface water",
                                                      "Purchased", "MHI", "Percent renter", "Percent non-Hispanic white"),
                                 omit = "Constant",
                                apply.ci = function(x) { 0 }))

cat(paste(gsub("\\(0.000, 0.000\\)", "", Compositestable), collapse = "\n"), "\n")

```

<br>
<br>

```{r table WQ, results='asis', echo=FALSE}
#WQ indicators
ecoliOR <- ecoli
ecoliOR$coefficients <- exp(ecoliOR$coefficients)
pvaluesecoli <- list(summary(ecoli)$coefficients[,4])

TTOR <- TT
TTOR$coefficients <- exp(TTOR$coefficients)
pvaluesTT <- list(summary(TT)$coefficients[,4])

MCLOR <- MCL
MCLOR$coefficients <- exp(MCLOR$coefficients)
pvaluesMCL <- list(summary(MCL)$coefficients[,4])

WQmodellist <- list(ecoli, ecoliOR, TT, TTOR, MCL, MCLOR)

WQtable <- capture.output(stargazer(WQmodellist,
                                 type = 'html', 
                                 column.labels = c("coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR"),
                                 model.numbers = FALSE, 
                                 dep.var.labels = c("E. coli", "Treatment technique violations", "MCL violations"), 
                                 ci=c(F,T, F, T, F, T, F, T), 
                                 star.cutoffs = c(0.05, 0.01, 0.001),
                                 notes="<span>&#42;&#42;&#42;</span>: p<0.001; <span>&#42;&#42;</span>: p<0.01;<span>&#42;</span>: p<0.05",
                                 notes.append = FALSE,
                                 p = c(pvaluesecoli, p.values2, pvaluesTT, p.values2,
                                       pvaluesMCL, p.values2),
                                 omit.stat = c("ll"),
                                 covariate.labels = c("Limited enfran.", "No enfran.",
                                                      "Population", "Surface water",
                                                      "Purchased", "MHI", "Percent renter", "Percent non-Hispanic white"),
                                 omit = "Constant",
                                 apply.ci = function(x) { 0 }))

cat(paste(gsub("\\(0.000, 0.000\\)", "", WQtable), collapse = "\n"), "\n")

```

<br>
<br>

```{r table affordability, results='asis', echo=FALSE}
#affordability indicators
MHIOR <- MHI
MHIOR$coefficients <- exp(MHIOR$coefficients)
pvaluesMHI <- list(summary(MHI)$coefficients[,4])

extremeOR <- extreme
extremeOR$coefficients <- exp(extremeOR$coefficients)
pvaluesextreme <- list(summary(extreme)$coefficients[,4])

covidOR <- covid
covidOR$coefficients <- exp(covidOR$coefficients)
pvaluescovid <- list(summary(covid)$coefficients[,4])

AFmodellist <- list(MHI, MHIOR, extreme, extremeOR, covid, covidOR)

AFtable <- capture.output(stargazer(AFmodellist,
                                 type = 'html', 
                                 column.labels = c("coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR"),
                                 model.numbers = FALSE, 
                                 dep.var.labels = c("Unaffordable bill", "Extreme bill",
                                                    "No arrearage relief"), 
                                 ci=c(F,T, F, T, F, T, F, T), 
                                 star.cutoffs = c(0.05, 0.01, 0.001),
                                 notes="<span>&#42;&#42;&#42;</span>: p<0.001; <span>&#42;&#42;</span>: p<0.01;<span>&#42;</span>: p<0.05",
                                 notes.append = FALSE,
                                 p = c(pvaluesMHI, p.values2, pvaluesextreme, p.values2, pvaluescovid, p.values2),
                                 omit.stat = c("ll"),
                                 covariate.labels = c("Limited enfran.", "No enfran.",
                                                      "Population", "Surface water",
                                                      "Purchased", "MHI", "Percent renter", "Percent non-Hispanic white"),
                                 omit = "Constant",
                                 apply.ci = function(x) { 0 }))

cat(paste(gsub("\\(0.000, 0.000\\)", "", AFtable), collapse = "\n"), "\n")

```

<br>
<br>

```{r table Accessibility, results='asis', echo=FALSE}
#Accessibility indicators
Single_sourceOR <- Single_source
Single_sourceOR$coefficients <- exp(Single_sourceOR$coefficients)
pvaluesSinglesource <- list(summary(Single_source)$coefficients[,4])

Sourcecapacity_violationsOR <- Sourcecapacity_violations
Sourcecapacity_violationsOR$coefficients <- exp(Sourcecapacity_violationsOR$coefficients)
pvaluesSourcecapacity_violations <- list(summary(Sourcecapacity_violations)$coefficients[,4])

Service_disOR <- Service_dis
Service_disOR$coefficients <- exp(Service_disOR$coefficients)
pvaluesService_dis <- list(summary(Service_dis)$coefficients[,4])

Accessmodellist <- list(Single_source, Single_sourceOR, Sourcecapacity_violations, Sourcecapacity_violationsOR, Service_dis, Service_disOR)

Accesstable <- capture.output(stargazer(Accessmodellist,
                                 type = 'html', 
                                 column.labels = c("coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR", 
                                                   "coefficient (se)", "OR"),
                                 model.numbers = FALSE, 
                                 dep.var.labels = c("Single source", "Source capacity violations", "Service disruptions"), 
                                 ci=c(F,T, F, T, F, T), 
                                 star.cutoffs = c(0.05, 0.01, 0.001),
                                 notes="<span>&#42;&#42;&#42;</span>: p<0.001; <span>&#42;&#42;</span>: p<0.01;<span>&#42;</span>: p<0.05",
                                 notes.append = FALSE,
                                 p = c(pvaluesSinglesource, p.values2, pvaluesSourcecapacity_violations, p.values2, pvaluesService_dis, p.values2),
                                 omit.stat = c("ll"),
                                 covariate.labels = c("Limited enfran.", "No enfran.",
                                                      "Population", "Surface water",
                                                      "Purchased", "MHI", "Percent renter", "Percent non-Hispanic white"),
                                 omit = "Constant",
                                 apply.ci = function(x) { 0 }))

cat(paste(gsub("\\(0.000, 0.000\\)", "", Accesstable), collapse = "\n"), "\n")
```

<br>
<br>

```{r descriptives for demofraphics, results='asis', echo=FALSE}

Data$Source_bi <- NA
Data$Source_bi <- ifelse(Data$Source == "GW", 1, Data$Source_bi)
Data$Source_bi <- ifelse(Data$Source == "SW", 0, Data$Source_bi)

Data_mini <- Data %>% select(enfranchisement_final, MHI, LN_POP, Source, Percent.renter, Percent.white, POPULATION, CALENVIRO_SCREEN_SCORE, FAILING, WQ_combined_any, Affordability_combined_any, Accessibility_combined_any, Percent.hispanicorlatino, Percent.black, Percent.asian, Source_bi)

Descriptives<- Data_mini %>%
  group_by(enfranchisement_final) %>% 
  summarize(meanPOP = mean(na.omit(POPULATION)),
            meanMHI = mean(na.omit(MHI)),
            meanPercRent = mean(na.omit(Percent.renter)),
            meanPercWhite = mean(na.omit(Percent.white)),
            meanCES = mean(na.omit(CALENVIRO_SCREEN_SCORE)),
            medianpop = median(na.omit(POPULATION)))

SIdescriptives <- Data_mini %>% 
  group_by(enfranchisement_final) %>% 
  select_if(is.numeric) %>% 
  gather(Variable, Value, -enfranchisement_final) %>% 
  group_by(Variable, enfranchisement_final) %>% 
  summarise(min = min(na.omit(Value)), 
            median = median(na.omit(Value)),
            mean = mean(na.omit(Value)), 
            max = max(na.omit(Value)),
            sd = sd(na.omit(Value))) %>% 
 ungroup()

SIdescriptives <- SIdescriptives %>% mutate_if(is.numeric, round, digits=3)
```

```{r dem sig tests, results='asis', echo=FALSE}
#ANOVA and Tukey HSD
MHI <- aov(MHI ~ enfranchisement_final, data = Data)
summary(MHI)
TukeyMHI <- TukeyHSD(MHI)
TukeyMHI_table <- as.data.frame(TukeyMHI$enfranchisement_final)

GW <- aov(Data$Source_bi ~ Data$enfranchisement_final)
summary(GW)
TukeyGW <- TukeyHSD(GW)
TukeyGW_table <- as.data.frame(TukeyGW$`Data$enfranchisement_final`)

Pop <- aov(Data$POPULATION ~ Data$enfranchisement_final)
summary(Pop)
TukeyPop <- TukeyHSD(Pop)
TukeyPop_table <- as.data.frame(TukeyPop$`Data$enfranchisement_final`)

Rent <- aov(Data$Percent.renter ~ Data$enfranchisement_final)
summary(Rent)
TukeyRent <- TukeyHSD(Rent)
TukeyRent_table <- as.data.frame(TukeyRent$`Data$enfranchisement_final`)

CalE <- aov(Data$CALENVIRO_SCREEN_SCORE ~ Data$enfranchisement_final)
summary(CalE)
TukeyCalE <- TukeyHSD(CalE)
TukeyCalE_table <- as.data.frame(TukeyCalE$`Data$enfranchisement_final`)

White <- aov(Data$Percent.white ~ Data$enfranchisement_final)
summary(White)
TukeyWhite <- TukeyHSD(White)
TukeyWhite_table <- as.data.frame(TukeyWhite$`Data$enfranchisement_final`)

AA <- aov(Percent.black ~ enfranchisement_final, data = Data)
summary(AA)
TukeyAA <- TukeyHSD(AA)
TukeyAA_table <- as.data.frame(TukeyAA$enfranchisement_final)

Lat <- aov(Data$Percent.hispanicorlatino ~ Data$enfranchisement_final)
summary(Lat)
TukeyLat <- TukeyHSD(Lat)
TukeyLat_table <- as.data.frame(TukeyLat$`Data$enfranchisement_final`)

As <- aov(Data$Percent.asian ~ Data$enfranchisement_final)
summary(As)
TukeyAs <- TukeyHSD(As)
TukeyAs_table <- as.data.frame(TukeyAs$`Data$enfranchisement_final`)

```